---
title: "5W Cleaning"
author: "Sean Ng"
date: "10/25/2021"
output: html_document
---

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(janitor)
library(readxl)
library(stringi) # stringi used for the rm_accents function
library(stringr)
library(timeDate)
library(pander)
```

```{r, reading in dataset, message=FALSE, warning = FALSE}
# reading in data
# act_raw<- read_excel("database activities 5W.xlsx", 
#                     sheet = "data", 
#                     skip = 1, 
#                     .name_repair = make_clean_names)


# Column types
act_col_types <- c("date", "date", "text", "text","text","text",
                   "text","text","text","text","text","text","text",
                   "numeric", "numeric", "text","text","text",
                   "text","text","text", "numeric", "text", 
                   "date", "date", "text", "numeric", "numeric",
                   "logical", "numeric", "numeric", "numeric",
                   "numeric", "numeric", "numeric", "numeric",
                   "numeric", "numeric", "numeric", "numeric",
                   "numeric", "numeric", "numeric")

# I think you're gonna have to live with the error messages
# since I don't think you can slice before read_excel()
# be aware of the columns you added that are in this dataset and 
# only in this dataset

act1<- read_excel("database activities 5W.xlsx", 
                     sheet = "data", 
                     skip = 1, 
                     col_types = act_col_types) %>% 
  clean_names() %>% 
  remove_empty() %>% 
  slice(-c(1)) %>% # removes the second row
  slice(-c(12055)) # removes the grand total

```

> Variables are renamed here -- all in lower case and in line with R syntax. A function to remove accent markes is then created and all location variables (and activity descriptions) are cleansed of accent marks and converted to uppercase. 

> This section handles the recoding of individual cells in various columns that have spelling errors and problems with standardisation. Checking arguments are at the bottom. Caution: this section has several overwrites. 

```{r, String Manipulations and Recoding}
# names vector, you actually might not need the clean_names() argument
act_names <- c("fecha_reportado_cluster",
               "mes_reportado",
               "codigo_de_proyecto",
               "nombre_del_proyecto",
               "organizacion_lider",
               "organizacion_implementadora",
               "estado",
               "pcode1",
               "municipio",
               "pcode2",
               "parroquia",
               "pcode3",
               "ubicacion",
               "latitud",
               "longitud",
               "actividad_full",
               "actividad_codigo",
               "actividad_desc",
               "actividad_codigo_desc",
               "act_covid",
               "unidad",
               "cantidad_del_mes",
               "recurrente_beneficiarios",
               "fecha_de_inicio",
               "fecha_prevista_finalizacion",
               "estatus_actividad",
               "total_beneficiarios",
               "check_beneficiaries",
               "t_f_ben",
               "percent_poblacion_indigena",
               "percent_personas_con_discapacidad",
               "m_0_3",
               "m_3_6",
               "m_7_12",
               "m_12_17",
               "m_18_19",
               "f_0_3",
               "f_3_6",
               "f_7_12",
               "f_12_17",
               "f_18_19",
               "f_mayores_de_19",
               "m_mayores_de_19")

# Renamed
names(act1) <- act_names

# function to remove accents
rm_accent <- function(colns){
  colns <- stri_trans_general(colns, "Latin-ASCII")
}

# removing accents
act1 <- act1 %>% 
  mutate(estado    = rm_accent(str_to_upper(estado)), # string to upper and rm_accent
         municipio = rm_accent(str_to_upper(municipio)),
         parroquia = rm_accent(str_to_upper(parroquia)),
         ubicacion = rm_accent(str_to_upper(ubicacion)),
         actividad_desc = rm_accent(str_to_upper(actividad_desc)))

# recoding recurrrent_beneficiarios
act1$recurrente_beneficiarios <- act1$recurrente_beneficiarios %>% 
  recode(no = FALSE, No = FALSE, Si = TRUE, si = TRUE, Sí = TRUE)

# overwriting NAs to be FALSE -- all NAs in recurrente_beneficiarios are FALSE
act1$recurrente_beneficiarios[is.na(act1$recurrente_beneficiarios)] <- FALSE

# recoding act_covid
act1$act_covid <- act1$act_covid %>% 
  recode(no = FALSE, No = FALSE, Si = TRUE, si = TRUE, Sí = TRUE)

# overwriting NAs to be FALSE -- all NAs in act_covid are treated as FALSE
act1$act_covid[is.na(act1$act_covid)] <- FALSE

# recoding estatus_actividad
act1$estatus_actividad <- act1$estatus_actividad %>% 
  str_replace_all(c("En ejecucion" = "ejecucion", 
                    "en ejecución" = "ejecucion", 
                    "en Ejecución" = "ejecucion",
                    "En ejecución" = "ejecucion",
                    "En Ejecución" = "ejecucion",
                    "finalizada" = "finalizada",
                    "Finalizada" = "finalizada"))


# create a new variable of last day of month only 
act1$mes_solo <- ceiling_date(act1$mes_reportado, "month") - days(1)

```

```{r checking-recode, eval=FALSE}
# checking, uncomment it if you need it
count(act2, recurrente_beneficiarios)
count(act2, act_covid)
count(act2, estatus_actividad)
```


> Cleaning locations -- you need to match to the locations dataset. You'll do this later since you don't have an updated locations file. 

```{r, Location Cleaning}
# checking individual variables
# act1[is.na(act1$pcode2),] %>%  glimpse()

# splitting into two datasets, one clean one dirty. we will bind_rows them 
# once adm_dirty is cleaned. 
adm_dirty <- act1 %>% 
  filter(is.na(estado) | is.na(pcode1) |
         is.na(municipio) | is.na(pcode2) |
         is.na(parroquia) | is.na(pcode3))

adm_clean <- act1 %>% 
  filter(!is.na(estado) & !is.na(pcode1) &
         !is.na(municipio) & !is.na(pcode2) &
         !is.na(parroquia) & !is.na(pcode3)) 

```

> This section reads in the locations reference dataset then uses it to clean `adm_dirty` and rewrites it to `adm_dirty1`. The `distinct()` call at the end is to ensure that no duplicates are included as the `left_join()` argument will product duplicates. After which, `adm_clean` and `adm_dirty1` are combined into `act2`. 
> you're going to need to include a new pcode3 for "todo municipio" and a new pcode2 for "todo estado"

```{r}
locations <- read_csv("locations.csv") %>%
  mutate(estado    = rm_accent(str_to_upper(estado)),
         municipio = rm_accent(str_to_upper(municipio)),
         parroquia = rm_accent(str_to_upper(parroquia)),
         ubicacion = rm_accent(str_to_upper(ubicacion)))

# for filling in missing values
adm_dirty1 <- adm_dirty %>% 
  left_join(locations, by = "ubicacion") %>% 
  mutate(estado    = coalesce(estado.x, estado.y),
         pcode1    = coalesce(pcode1.x, pcode1.y),
         municipio = coalesce(municipio.x, municipio.y),
         pcode2    = coalesce(pcode2.x, pcode2.y),
         parroquia = coalesce(parroquia.x, parroquia.y),
         pcode3    = coalesce(pcode3.x, pcode3.y)) %>% 
  select(-estado.x, -estado.y, -pcode1.x, -pcode1.y,
         -municipio.x, -municipio.y, -pcode2.x, -pcode2.y,
         -parroquia.x, -parroquia.y, -pcode3.x, -pcode3.y) %>% 
  distinct()

# bind_rows() does not care about column sequence 
# but put adm_clean first in the bind_rows() as the sequence is important later
act2 <- bind_rows(adm_clean, adm_dirty1)

# checking that the number of beneficiaries has stayed the same
sum(act1$total_beneficiarios, na.rm = TRUE)
sum(act2$total_beneficiarios, na.rm = TRUE)

sum(is.na(act1$total_beneficiarios))
```

> This is to check that `total_beneficiarios` is equal to the sum of all disaggregated columns. First, we mutate a new column by summing all the disaggregated beneficiaries and calling it `total_ben_check`. The next function prints the differences between `total_ben_check` and `total_beneficiarios`. Finally, we mutate a new column in the dataset called `no_esp_ben` or beneficiarios no especificados to collect the differences and ensure that we do not lose them. 

```{r, bencheck}

# mutating bencheck and unspecified beneficiaries columns 
act2 <- act2 %>% 
  mutate(total_ben_check = select(., m_0_3:m_mayores_de_19) %>%  rowSums(na.rm = TRUE), 
         no_esp_ben      = round(total_beneficiarios) - round(total_ben_check)) 

# printing the bencheck -- I think you need a better option than pander()
# which provides marginal totals
# the round is necessary due to floating point arithmetic as some values are
# close enough to 0 to be ignored by us but not by R. 

panderOptions('table.split.table', Inf)

# only pander() if you want to read the html
act2 %>% 
  filter(round(total_beneficiarios) != round(total_ben_check)) %>% 
  select(organizacion_implementadora, estado, municipio, parroquia, actividad_desc, 
         ubicacion, total_ben_check, total_beneficiarios, no_esp_ben) 
# %>% pander()

```

> This section pivots the dataset longer into tidy data which, with one observation per instance of age and sex; this is necessary for easy plotting and additional operations. Additionally, rows with no beneficiaries are filtered out. 

> The dataset `u_ben` is created by taking the highest number of beneficiaries by age and sex group by location. The columns are then filtered down to create a more usable dataset. We will use this dataset for any operations related to unique beneficiaries (individuals), including comparisons with census data to determine reach and coverage. Please note that not all activities will be represented in this dataset as only the max of beneficiaries per disaggregation group per location was selected.  

```{r, Unique Benficiaries and Pivoting Longer}

# pivoting longer and creating new dataframe
  
u_ben <- act2 %>% 
  select(-t_f_ben, -total_ben_check) %>% # remember to drop them
  relocate(mes_solo) %>% # moves mes_solo to the first variable
  pivot_longer(m_0_3:no_esp_ben, 
               names_to = "desagregacion", values_to = "beneficiarios") %>% 
  filter(recurrente_beneficiarios == FALSE) %>%
  filter(beneficiarios != 0) %>% # empty cells are 0 in the 5W table
  group_by(ubicacion, desagregacion) %>% 
  slice(which.max(beneficiarios)) %>% 
  ungroup() %>% 
  select(mes_solo, estado, pcode1, municipio, pcode2,parroquia, pcode3, 
         ubicacion, latitud, longitud,
         desagregacion, beneficiarios)

# writing csv of u_ben with datestamp
now <- Sys.time()
filename_u_ben <- paste0(format(now, "%y%m%d_%H%M_"), "u_ben.csv")
write_csv(u_ben, filename_u_ben)
  
```

> The dataset `act_ben` is for activity totals. Only the rows marked as `recurrente_beneficiarios == FALSE` have been selected. The first instance of recurring beneficiaires is marked FALSE, with all subsequent entries being marked TRUE. Similar to `u_ben`, it has also been pivoted longer into tidy data. 

> Please not that we can only assure that there no duplications within locations as specific locations might have been targetted by more than one activity. For geographic operations, please use `u_ben`.


```{r, Beneficiairies per activity}

# This is the dataset for beneficiaries
act_ben <- act2 %>%
  select(-t_f_ben, -total_ben_check) %>% # remember to drop them
  relocate(mes_solo) %>% # moves mes_solo to the first variable
  filter(recurrente_beneficiarios == FALSE) %>%
  pivot_longer(m_0_3:no_esp_ben, 
               names_to = "desagregacion", values_to = "beneficiarios") %>% 
  filter(beneficiarios != 0) %>% # empty cells are 0 in the 5W table
  group_by(ubicacion, desagregacion) %>% 
  select(mes_solo, nombre_del_proyecto, organizacion_implementadora, 
         estado, pcode1, municipio, pcode2, parroquia, pcode3, 
         ubicacion, latitud, longitud,
         actividad_codigo, actividad_desc, act_covid,
         desagregacion, beneficiarios)

# writing csv act_ben with datestamp
now <- Sys.time()
filename_act_ben <- paste0(format(now, "%y%m%d_%H%M_"), "act_ben.csv")
write_csv(act_ben, filename_act_ben)

# comparing u_ben and act_ben
sum(act_ben$beneficiarios)
sum(u_ben$beneficiarios)
```

