---
title: "Cleaning and Reporting on Education 5W data from Venezuela"
author: "Sean Ng"
date: "12 November, 2021"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: false
    collapsed: false
    code_download: true
    theme: readable
---

# A. Cleaning the 5W data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 9.5)
library(tidyverse)
library(lubridate)
library(stringr)
library(stringi)
library(pander)
library(scales)
library(forcats)
library(janitor)
library(readxl)
library(ggmap)
library(sf)
library(patchwork)
library(maptools)
library(ggforce)
library(ggthemes)
library(plotly)
library(kableExtra)
library(DT)
library(tinytex)

# tables all in one row
panderOptions('table.split.table', Inf)

# thousands separator
panderOptions("big.mark", ",")

```


> This first section details an example of the code necessary for cleaning a 5W dataset in R. 5W data is typically messy and in wide format. Additionally, data wrangling is time consuming and an inefficient use of a specialist's time. The code presented below intends to demonstrate that this monthly (or sometimes more frequent) occurence can be greatly automated, allowing for more timely and accurate analysis as well as creating the space for more tailored reporting.  

&nbsp;

## 1. Reading in the dataset

```{r reading-in-xlsx, echo=TRUE}
# specifying column types 
act_col_types <- c("date", "date", "text", "text","text","text",
                   "text","text","text","text","text","text","text",
                   "numeric", "numeric", "text","text","text",
                   "text","text","text", "numeric", "text", 
                   "date", "date", "text", "numeric", "numeric",
                   "logical", "numeric", "numeric", "numeric",
                   "numeric", "numeric", "numeric", "numeric",
                   "numeric", "numeric", "numeric", "numeric",
                   "numeric", "numeric", "numeric")

# you're gonna have to live with the error messages,
# since you can't slice before read_excel() argument

act1<- read_excel("database activities 5W.xlsx", 
                     sheet = "data", 
                     skip = 1, 
                     col_types = act_col_types) %>% 
  clean_names() %>% 
  remove_empty() %>% 
  slice(-c(1)) %>% # removes the second row
  slice(-c(12055)) # removes the grand total

```

### 1a. Renaming Columns 

> A `glimpse()` of the renamed dataset:

```{r}
act_names <- c("fecha_reportado_cluster",
               "mes_reportado",
               "codigo_de_proyecto",
               "nombre_del_proyecto",
               "organizacion_lider",
               "organizacion_implementadora",
               "estado",
               "pcode1",
               "municipio",
               "pcode2",
               "parroquia",
               "pcode3",
               "ubicacion",
               "latitud",
               "longitud",
               "actividad_full",
               "actividad_codigo",
               "actividad_desc",
               "actividad_codigo_desc",
               "act_covid",
               "unidad",
               "cantidad_del_mes",
               "recurrente_beneficiarios",
               "fecha_de_inicio",
               "fecha_prevista_finalizacion",
               "estatus_actividad",
               "total_beneficiarios",
               "check_beneficiaries",
               "t_f_ben",
               "percent_poblacion_indigena",
               "percent_personas_con_discapacidad",
               "m_0_3",
               "m_3_6",
               "m_7_12",
               "m_12_17",
               "m_18_19",
               "f_0_3",
               "f_3_6",
               "f_7_12",
               "f_12_17",
               "f_18_19",
               "f_mayores_de_19",
               "m_mayores_de_19")

# renamed 
names(act1) <- act_names

glimpse(act1)
```

&nbsp;

## 2. String Manipulation and Recoding 

### 2a. Removing accents and standardising case

> Whilst it is true when some people say that if you remove tildes and other diacritical marks, it it no longer becomes Spanish. However, it must be said that implementing partners don't care and have included a wide variety of spellings. I would say that remove all accent marks is necessary to continue working with the dataset. 

```{r remove-accents, echo=TRUE}
# function to remove accents
rm_accent <- function(colns){
  colns <- stri_trans_general(colns, "Latin-ASCII")
}

# removing accents
# and str_to_upper() since R is case-sensitive
act1 <- act1 %>% 
  mutate(estado    = rm_accent(str_to_upper(estado)), 
         municipio = rm_accent(str_to_upper(municipio)),
         parroquia = rm_accent(str_to_upper(parroquia)),
         ubicacion = rm_accent(str_to_upper(ubicacion)),
         actividad_desc = rm_accent(str_to_upper(actividad_desc)))

```

### 2b. Recoding erroneous data

> These should just be dropdown menus; partners should also be informed that blanks will be treated as FALSE. Additionally, a new date column just with the month is mutated, as the `mes_reportado` column has some months with multiple reporting dates. 

```{r recoding, echo=TRUE}
# recoding 
act1 <- act1 %>% 
  mutate(recurrente_beneficiarios = 
           recode(recurrente_beneficiarios, 
            "no" = FALSE, "No" = FALSE, "Si" = TRUE, "si" = TRUE, "sí" = TRUE, "Sí" = TRUE),
         act_covid = recode(act_covid, 
            "no" = FALSE, "No" = FALSE, "Si" = TRUE, "si" = TRUE, "sí" = TRUE, "Sí" = TRUE)) %>% 
  replace_na(list(recurrente_beneficiarios = FALSE, act_covid = FALSE)) 

# recoding estatus_actividad
act1$estatus_actividad <- act1$estatus_actividad %>% 
  str_replace_all(c("En ejecucion" = "ejecucion", 
                    "en ejecución" = "ejecucion", 
                    "en Ejecución" = "ejecucion",
                    "En ejecución" = "ejecucion",
                    "En Ejecución" = "ejecucion",
                    "finalizada" = "finalizada",
                    "Finalizada" = "finalizada"))

# create a new variable of last day of month only 
act1$mes_solo <- ceiling_date(act1$mes_reportado, "month") - days(1)
```

&nbsp;

## 3. Cleaning locations 

### 3a. Read in `locations.csv`

> And check if you need to update it. 

```{r read-locations, echo=TRUE}
locations <- read_csv("locations.csv") %>%
  mutate(estado    = rm_accent(str_to_upper(estado)), # just to make sure 
         municipio = rm_accent(str_to_upper(municipio)),
         parroquia = rm_accent(str_to_upper(parroquia)),
         ubicacion = rm_accent(str_to_upper(ubicacion)))
```

> Maybe we should add new pcodes for "todo municipio" and  "todo estado"; will they be abused?

```{r, locations-check, echo=TRUE}
# see if you need to update locations 
# new locations not in locations.csv will be caught by the anti-join()
 locations_add <- act1 %>% 
   select(estado, pcode1, municipio, pcode2, parroquia, pcode3, ubicacion,
          latitud, longitud) %>% 
   distinct() %>% 
   anti_join(locations, by = "ubicacion")
```


### 3b. Splitting the dataset into clean and dirty

```{r split-clean-dirty, echo=TRUE}
# splitting into two datasets, one clean one dirty. 
adm_dirty <- act1 %>% 
  filter(is.na(estado) | is.na(pcode1) |
         is.na(municipio) | is.na(pcode2) |
         is.na(parroquia) | is.na(pcode3))

adm_clean <- act1 %>% 
  filter(!is.na(estado) & !is.na(pcode1) &
         !is.na(municipio) & !is.na(pcode2) &
         !is.na(parroquia) & !is.na(pcode3)) 

```

### 3c. Fill in missing values with `coalesce()`

> The locations reference dataset is used to clean `adm_dirty` and rewrite it. The `distinct()` call at the end is to ensure that no duplicates are included as the `left_join()` argument will produce duplicates (still not sure how to control this). After which, `adm_clean` and `adm_dirty` are combined into `act2`.

```{r fill in missing values, echo=TRUE}
# filling in missing values
adm_dirty <- adm_dirty %>% 
  left_join(locations, by = "ubicacion") %>% 
  mutate(estado    = coalesce(estado.x, estado.y),
         pcode1    = coalesce(pcode1.x, pcode1.y),
         municipio = coalesce(municipio.x, municipio.y),
         pcode2    = coalesce(pcode2.x, pcode2.y),
         parroquia = coalesce(parroquia.x, parroquia.y),
         pcode3    = coalesce(pcode3.x, pcode3.y)) %>% 
  select(-estado.x, -estado.y, -pcode1.x, -pcode1.y,
         -municipio.x, -municipio.y, -pcode2.x, -pcode2.y,
         -parroquia.x, -parroquia.y, -pcode3.x, -pcode3.y) %>% 
  distinct() # removing duplicate rows from the join 

# bind_rows() does not care about column sequence 
# but put adm_clean first so that the original order is preserved
act2 <- bind_rows(adm_clean, adm_dirty)

```

### 3d. Error checking the join
> printing total beneficiaries for `act1` and `act2`, followed by `nrow()` for both. 

```{r join-error-check}
# error checking
sum(act1$total_beneficiarios, na.rm = TRUE)
sum(act2$total_beneficiarios, na.rm = TRUE)
nrow(act1)
nrow(act2)
```

&nbsp;

## 4. Beneficiary disaggregation 
> This is to check that `total_beneficiarios` is equal to the sum of all disaggregated columns. First, we mutate a new column by summing all the disaggregated beneficiaries and calling it `total_ben_check`. Then, we mutate a new column in thr dataset called `no_esp_ben` or beneficiarios no especificados so that we maintain the un-disaggregated beneficiaries without polluting the rest of the dataset.

```{r bencheck-and-no-esp-ben, echo=TRUE }
# mutating bencheck and unspecified beneficiaries columns 
act2 <- act2 %>% 
  mutate(total_ben_check = select(., m_0_3:m_mayores_de_19) %>%  rowSums(na.rm = TRUE), 
         no_esp_ben      = round(total_beneficiarios) - round(total_ben_check)) 

```

> Next, we print rows where the sum of disaggregated beneficiaries (`total_ben_check`) do not match the totals reported in the `total_beneficiarios` column so that we can return to the relevant parties for follow up. 

```{r printing-bencheck}
# printing the bencheck
# the round is necessary due to floating point arithmetic as some values are
# close enough to 0 to be ignored by us but not by R. 
# only pander() if you want to read the html
no_esp <- act2 %>% 
  filter(round(total_beneficiarios) != round(total_ben_check)) %>% 
  select(organizacion_implementadora, actividad_desc, 
         ubicacion, total_ben_check, total_beneficiarios, no_esp_ben)

kable(no_esp, caption = "Undisaggregated beneficiaries") %>%
  kable_styling(bootstrap_options = "bordered", font_size = 12)
```

&nbsp;

## 5. Outputs

### 5a. Cleaned wide version
```{r datestamp, echo=TRUE}
# timestamp for all the files to be written below
now <- Sys.time()
```

> For submissions to OCHA, this one aligns with their format

```{r cleaned-wide, echo=TRUE, eval=FALSE}
# turn on eval if you want the file
filename_wide <- paste0(format(now, "%y%m%d_%H%M_"), "5W_wide.csv")
write_csv(act2, filename_wide)
```

### 5b. Unique beneficiaries

> This section pivots the dataset longer into tidy data, with one observation per instance of age and sex; this is necessary for easy plotting and analysis. Additionally, rows with no beneficiaries are filtered out. 

> The dataset `u_ben` is created by taking the highest number of beneficiaries by age and sex group by location of all non-recurrent beneficiaries. The columns are then filtered down to create a more usable dataset. We will use this dataset for any operations related to unique beneficiaries (individuals), including comparisons with census data to determine reach and coverage as well as geographic analysis. We drop activity information from this dataset as it only contains the max of per disaggregation group per location anyway. 

```{r unique-benficiaries, echo=TRUE}
# pivoting longer and creating new dataframe
u_ben <- act2 %>% 
  select(-t_f_ben, -total_ben_check) %>% # remember to drop them
  relocate(mes_solo) %>% # moves mes_solo to the first variable
  pivot_longer(m_0_3:no_esp_ben, 
               names_to = "desagregacion", values_to = "beneficiarios") %>% 
  filter(recurrente_beneficiarios == FALSE) %>%
  filter(beneficiarios != 0) %>% # empty cells are 0 in the 5W table
  group_by(ubicacion, desagregacion) %>% 
  slice(which.max(beneficiarios)) %>% 
  ungroup() %>% 
  select(mes_solo, estado, pcode1, municipio, pcode2,parroquia, pcode3, 
         ubicacion, latitud, longitud,
         desagregacion, beneficiarios)
```

```{r writing-u-ben-csv, echo=TRUE, eval=FALSE}
# writing csv of u_ben with datestamp
# turn on eval if you want the file 
filename_u_ben <- paste0(format(now, "%y%m%d_%H%M_"), "u_ben.csv")
write_csv(u_ben, filename_u_ben)
```


### 5c. Beneficiaries by activity

> The dataset `act_ben` is for activity-wise analysis. Only the rows marked as `recurrente_beneficiarios == FALSE` have been selected. The first instance of recurring beneficiaires is marked FALSE, with all subsequent entries being marked TRUE. Similar to `u_ben`, it has also been pivoted longer into tidy data. 

> Please not that there is double counting in this dataset as specific beneficiaries might been reached by more than one activity. For geographic analysis, please use `u_ben`.

```{r beneficiaries-by-activity, echo=TRUE}
# This is the dataset for beneficiaries
act_ben <- act2 %>%
  select(-t_f_ben, -total_ben_check) %>% # remember to drop them
  relocate(mes_solo) %>% # moves mes_solo to the first variable
  filter(recurrente_beneficiarios == FALSE) %>%
  pivot_longer(m_0_3:no_esp_ben, 
               names_to = "desagregacion", values_to = "beneficiarios") %>% 
  filter(beneficiarios != 0) %>% # empty cells are 0 in the 5W table
  group_by(ubicacion, desagregacion) %>% 
  select(mes_solo, nombre_del_proyecto, organizacion_implementadora, 
         estado, pcode1, municipio, pcode2, parroquia, pcode3, 
         ubicacion, latitud, longitud,
         actividad_codigo, actividad_desc, act_covid,
         desagregacion, beneficiarios)

```


```{r writing-act-ben-csv, echo=TRUE, eval=FALSE}
# writing csv act_ben with datestamp
# turn on eval if you want the file 
filename_act_ben <- paste0(format(now, "%y%m%d_%H%M_"), "act_ben.csv")
write_csv(act_ben, filename_act_ben)
```

### 5d. Comparing `u_ben`, `act_ben` and difference

> Please note that this still has that radio messaging activity lumped in with the totals. 

```{r comparing}

(format(sum(u_ben$beneficiarios), big.mark = ","))
(format(sum(act_ben$beneficiarios), big.mark = ","))
(format(sum(act_ben$beneficiarios) - sum(u_ben$beneficiarios), big.mark = ","))
```

&nbsp;

# B. Reporting on the 5W data 

```{r report-setup, include=FALSE}
# read in census reference dataset
cen_ref <- read_excel("census_data_20191122.xlsx", sheet = "data") %>% 
  clean_names() %>%
  select(estado, pcode1, municipio, pcode2, parroquia, pcode3, field_office,
         poblacion_total_2011, x_2019_poblacion_parroquial_total,
         ham_2019_ambitos_ge, ham_2019_xx_pobreza_env_por_parroquia, 
         ham_2019_xx_poblacion_pobre_por_parroquia,
         poblacion_infantil_menor_de_12_anos, poblacion_adolescentes_de_12_a_17_anos,
         poblacion_de_18_anos_y_mas, 
         poblacion_urbana_percent, area_km2, densidad_poblacional_ppl_km2,
         matricula_2017_educacion_inicial, matricula_2017_educacion_primaria, 
         matricula_2017_educacion_media)

# filtering out PROMOCION MENSAJES CLAVES PARA LA COMUNIDAD ESCOLAR
CA105 <- act_ben %>% filter(actividad_codigo == "CA1.05")

# creating datasets without CA1.05 and
# for act_ben, recoding UNICEF and mutating short activity descriptions
act_ben2 <- act_ben %>% 
  filter(actividad_codigo != "CA1.05") %>% 
  mutate(organizacion_implementadora = 
           str_replace_all(organizacion_implementadora, 
                           "UNICEF - Fondo de las Naciones Unidas para la Infancia",
                           "UNICEF")) %>% 
  mutate(actividad_short = recode(actividad_codigo, 
                             CA1.01 = "1.01_reinsercion educativa",
                             CA1.02 = "1.02_nivelacion y capacitacion",
                             CA1.03 = "1.03_kits escolares",
                             CA1.04 = "1.04_alimentacion escolar",
                             CA1.05 = "1.05_promocion mensajes claves",
                             CA1.06 = "1.06_apoyo psicoeducativo",
                             CA1.07 = "1.07_actividades recreativas",
                             CA2.08 = "2.08_educacion a distancia",
                             CA2.09 = "2.09_formacion docentes",
                             CA2.10 = "2.10_becas docentes"))
                    
u_ben2 <- u_ben %>% filter(!str_detect(.$ubicacion, "^RADIO"))
```

&nbsp;

> This is an entirely automated report -- all charts and tables, as well as all figures within the report have been generated from the data, with no manual input. This report makes use of the outputs of the `5W_cleaning` section above. This set of 5W data pertains to the Education Cluster in Venezuela and has had partner information removed. Code will not be shown in this section -- you can download the Rmd at the top-left corner of this doucument.  

&nbsp;

## 1. Summary of beneficiaries by activity, with sex ratio

```{r summary-TABLE-activity-and-sex}

# as a note, i don't think you can put an str_detect inside a summarise
act_ben2 %>% 
  mutate(sex_ben = ifelse(str_detect(desagregacion, "^m"), "male",
                          ifelse(str_detect(desagregacion, "^f"), "female", "no_esp"))) %>% 
  mutate(actividad = actividad_desc) %>% 
  group_by(actividad) %>% 
  summarise(total = sum(beneficiarios, na.rm = TRUE),
            male = sum((beneficiarios[sex_ben == "male"])),
            female = sum((beneficiarios[sex_ben == "female"])),
            sex_ratio = round((male/female), digits = 2)) %>% 
  mutate(percent_of_total = round((total / sum(total) * 100), digits = 2)) %>% 
  relocate(percent_of_total, .after = total) %>% 
  arrange(desc(total)) %>% 
  pander(big.mark = ",", style = "rmarkdown")

```

> A total of __`r format(round(sum(u_ben2$beneficiarios)), big.mark = ",")`__ individuals have been reached to date. In terms of frequencies (inclusive of double counting), __`r format(sum(act_ben2$beneficiarios), big.mark = ",")`__ have been reached. 

> Additionally, the __`r format(sum(CA105$beneficiarios), big.mark = ",")`__ beneficiary frequencies reached by the activity PROMOCION MENSAJES CLAVES PARA LA COMUNIDAD ESCOLAR have been removed from the totals in this report as the activity consists of solely radio messaging. 

&nbsp;

## 2. Beneficiaries by age group
*figures are unique beneficiaries/individuals*

```{r PLOT-beneficiaries-by-age}
# to reorder the plot
ordvec <- c("0-3", "3-6", "7-12", "12-17", "18-19", "mayor_que_19")

# df for the plot
age_u_ben <- u_ben2 %>% # important to use u_ben here 
  mutate(age_ben = case_when(
    str_detect(desagregacion,"0_3$") ~ "0-3",
    str_detect(desagregacion,"3_6$") ~ "3-6",
    str_detect(desagregacion, "7_12$") ~ "7-12", 
    str_detect(desagregacion, "12_17$") ~ "12-17",
    str_detect(desagregacion, "18_19$") ~ "18-19",
    str_detect(desagregacion, "mayores_de_19$") ~ "mayor_que_19", 
    TRUE ~ "no_esp")) %>%
  filter(age_ben != "no_esp") %>% 
  group_by(age_ben) %>% 
  summarise(beneficiarios = sum(beneficiarios))

ggplot(data = age_u_ben, aes(x = fct_reorder(age_ben, ordvec), y = beneficiarios)) +
  geom_col(fill = "cornflowerblue") +
  geom_text(aes(label = scales::comma(beneficiarios)), vjust = -0.25, size = 4) +
  scale_y_continuous(labels = unit_format(unit = "K", scale = 1e-3, sep = "")) +
  xlab("Grupos por Edad") + ylab("Beneficiarios") +
  theme(axis.text = element_text(size = 10))

```

&nbsp;

### 2a. Beneficiaries by age compared to 2017 Enrollment

```{r REF-matric-ben}
# reference df for the table and text below, restructuring enrollment data from cen_ref
matric_ben <- cen_ref %>% 
  select(matricula_2017_educacion_inicial, matricula_2017_educacion_primaria, 
         matricula_2017_educacion_media) %>%
  rename("3-6" = matricula_2017_educacion_inicial,
         "7-12" = matricula_2017_educacion_primaria,
         "12-17" = matricula_2017_educacion_media) %>% 
  pivot_longer(everything(), names_to = "age_ben", values_to = "matricula2017", 
               values_drop_na = TRUE) %>% 
  group_by(age_ben) %>% 
  summarise(matricula2017 = sum(matricula2017)) %>%
  left_join(age_u_ben) %>% 
  mutate(percent_total = beneficiarios/matricula2017 * 100) %>% 
  relocate(beneficiarios, .after = age_ben) %>% 
  rename(`Edad grupo` = age_ben) %>% 
  slice(2,3,1)  # rearranging rows
  
```

> With reference to the 2017 Matricula dataset, we can see that the Education programme has reached an overall average of __`r round(sum(matric_ben$beneficiarios)/(sum(matric_ben$matricula2017))*100)`%__ of schoolgoing children aged 3-17 nationwide. Children aged 3-17 consitute __`r round(sum(matric_ben$beneficiarios)/(sum(u_ben2$beneficiarios))*100)`%__ of all UNICEF beneficiaries. 

```{r TABLE-matric-ben}
pander(matric_ben, big.mark = ",", style = "rmarkdown")
```

&nbsp;

## 3. Changes since previous month

```{r REF-for-text-cumulative-datasets}

# creating reference dfs for the text and plot below
cum_u_ben <- u_ben2 %>% 
  mutate(mes_solo = as.Date(mes_solo)) %>% 
  arrange(mes_solo) %>% 
  group_by(mes_solo) %>% 
  summarise(sum_ben = sum(beneficiarios)) %>% 
  mutate(cumulative = cumsum(sum_ben))

cum_act_ben <- act_ben2 %>% 
  mutate(mes_solo = as.Date(mes_solo)) %>% 
  arrange(mes_solo) %>% 
  group_by(mes_solo) %>% 
  summarise(sum_ben = sum(beneficiarios)) %>% 
  mutate(cumulative = cumsum(sum_ben))
```

> The number of individuals reached has increased by __`r format(cum_u_ben[11,3] - cum_u_ben[10,3], big.mark = ",")`__ in the past month, reaching a total of __`r format(round(sum(u_ben2$beneficiarios)), big.mark = ",")`__. The number of beneficiary frequencies reached has increased by __`r format(cum_act_ben[11,3] - cum_act_ben[10,3], big.mark = ",")`__ in the same period, reaching a total of __`r format(sum(cum_act_ben[11,3]), big.mark = ",")`__.

```{r  line-PLOT-cumulative-beneficiaries}
u_ben2 %>% 
  mutate(mes_solo = as.Date(mes_solo)) %>% 
  arrange(mes_solo) %>% 
  group_by(mes_solo) %>% 
  summarise(sum_ben = sum(beneficiarios)) %>% 
  mutate(cumulative = cumsum(sum_ben)) %>% 
  ggplot(aes(x = mes_solo, y = cumulative)) +
  geom_line() +
  geom_text(aes(label = scales::comma(cumulative)), vjust = -0.25, hjust = 1, size = 3) +
  scale_x_date(date_labels = "%b-%Y", date_breaks = "1 month") +
  scale_y_continuous(labels = scales::unit_format(unit = "K", scale = 1e-3, sep = ""),
                     breaks = seq(0,950000, by = 200000)) + # what is this breaks doing?
  xlab("Mes Reportado") + ylab("Beneficiarios (cumulativo)")
```

&nbsp;

### 3a. Progress by activity by month
*mouse over to see details*


```{r line-PLOTLY-progress-by-activity, fig.width=9}
act_lines <- act_ben2 %>% 
  mutate(mes_solo = as.Date(mes_solo)) %>% 
  group_by(actividad_short, mes_solo) %>% 
  summarise(cumulative = sum(beneficiarios)) %>% 
  mutate(cumulative = round(cumsum(cumulative))) %>% 
  ggplot(aes(x = mes_solo, y = cumulative, colour = actividad_short)) +
  geom_line() +
  geom_point(size = 0.7) +
  scale_x_date(date_labels = "%b-%Y", date_breaks = "1 month") +
  scale_y_continuous(labels = unit_format(unit = "K", scale = 1e-3, sep = ""), 
                     limits = c(0, 500000)) +
  xlab("") + ylab("Beneficiarios") +
  scale_colour_discrete(name = "Actividad") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.text = element_text(size = 6)) +
  guides(fill = guide_legend(nrow = 2, byrow = TRUE))

ggplotly(act_lines, tooltip = c("x", "y", "colour")) %>% 
  layout(showlegend = TRUE, legend = list(font = list(size = 6))) %>% 
   config(displayModeBar = FALSE)
```
> Progress in recent months has largely been due to the distribution of education kits and distance learning. 

&nbsp;

## 4. Summaries by geography

### 4a. Beneficiaries by state

```{r PLOT-beneficiairies-by-state}
u_ben2 %>% 
  group_by(estado) %>% 
  summarise(beneficiarios = sum(beneficiarios)) %>%
  ggplot(aes(x = fct_reorder(estado, beneficiarios, .desc = TRUE), 
             y = beneficiarios)) +
  geom_col(fill = "cornflowerblue") +
  ylim(0, 850000) +
  geom_text(aes(label = scales::comma(beneficiarios, accuracy = 1)), 
            vjust = -0.25, hjust = 0.05, size = 3, angle = 30) +
  scale_y_continuous(labels = unit_format(unit = "K", scale = 1e-3, sep = ""), 
                     limits = c(0, 200000)) +
  xlab("") + ylab("Beneficiarios") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

&nbsp;

```{r REF-for-ubicacion-text}
mira_zul <- u_ben2 %>% filter(estado == "MIRANDA" | 
                              estado == "ZULIA")
```

### 4b. Number of schools by state
> A total of __`r format(n_distinct(u_ben2$ubicacion), big.mark = ",")`__ schools have been reached by UNICEF; __`r round((n_distinct(mira_zul$ubicacion)) / (n_distinct(u_ben2$ubicacion)) * 100)`%__ are from Miranda and Zulia alone. 

```{r PLOT-of-ubicacion-by-state}
u_ben2 %>% 
  group_by(estado) %>% 
  distinct(ubicacion) %>% 
  summarise(ubicacion = n()) %>% 
  ggplot(aes(x = fct_reorder(estado, ubicacion, .desc = TRUE), y = ubicacion)) +
  geom_col(fill = "cornflowerblue") +
  geom_text(aes(label = round(ubicacion)), vjust = -0.25, size = 3) + 
  xlab("") + ylab("Numero de Escuelas") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

&nbsp;

```{r REF-for-all_mun}
# creating the all_mun df
# all_mun is one obs per pcode2
all_mun <- u_ben2 %>% 
  mutate(is_matric = case_when( # for beneficiaries aged 3-17
        str_detect(desagregacion,"3_6$") ~ "matric",
        str_detect(desagregacion,"7_12$") ~ "matric",
        str_detect(desagregacion,"12_17$") ~ "matric",
        TRUE ~ "non_matric")) %>% 
  mutate(beneficiarios_3_17 = ifelse(is_matric == "matric", beneficiarios, 0)) %>% 
  group_by(estado, municipio, pcode2) %>% 
  summarise(beneficiarios = round(sum(beneficiarios)),
            beneficiarios_3_17 = round(sum(beneficiarios_3_17))) %>%
  ungroup() %>% 
  left_join(act_ben %>% # bringing activity counts into u_ben
            group_by(pcode2) %>% 
            summarise(act_types = n_distinct(actividad_codigo))) %>% 
  right_join(cen_ref %>% # for bringing matricula2017 into all_mun
            pivot_longer(cols = c("matricula_2017_educacion_inicial",
                              "matricula_2017_educacion_primaria", 
                              "matricula_2017_educacion_media"), 
                     names_to = "age_ben", values_to = "matricula2017", 
                     values_drop_na = TRUE) %>% 
            select(pcode2, age_ben, matricula2017) %>% 
            group_by(pcode2) %>% 
            summarise(matricula2017 = sum(matricula2017))) %>% 
  mutate(coverage_percent = round(beneficiarios_3_17 / matricula2017 * 100)) %>% 
  mutate(estado = fct_reorder(estado, beneficiarios, .desc = TRUE)) %>% 
  arrange(desc(beneficiarios))
```

### 4c. Scatterplot of municipalities
*logarithmic scale; larger points indicate more beneficiaries reached, darker blues indicate more activity types* 

*mouse over municipalities to see beneficiaries and distinct activities*

> A total of __`r n_distinct(u_ben2$pcode2)`__ municipalities were reached by the UNICEF Education programme.

```{r scatter-PLOTLY-of-municipalities-by-state, fig.width=9}

all_mun_points <- all_mun %>% 
  ggplot(aes(text = municipio, x = estado, y = beneficiarios, 
             colour = act_types)) +
  geom_point(aes(size = beneficiarios), alpha = 0.9) +
  scale_colour_gradient(low = "azure4", high = "darkblue") + #find a better colour?
  scale_y_continuous(trans = "log10", 
                     labels = comma) +
  xlab("") + ylab("Beneficiarios Alcanzados por Municipio") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none") 

ggplotly(all_mun_points, tooltip = c("text", "y", "colour")) %>% 
  config(displayModeBar = FALSE)

```

&nbsp;

### 4d. Top 10 municipalities by reach and coverage

```{r municipalities-top-TABLE}
# top municipalities table
top_mun <- all_mun %>% select(estado, municipio, beneficiarios) %>% slice(1:10)

# try to make a table of the most oversubscribed municipalities
top_cov <- all_mun %>% 
  arrange(desc(coverage_percent)) %>% 
  select(estado, municipio, coverage_percent) %>% slice(1:10)

# two tables side by side
kable(top_mun, caption = "by beneficiaries") %>%
  kable_styling(bootstrap_options = "bordered", full_width = FALSE, 
                position = "float_left", font_size = 12)
kable(top_cov, caption = "by coverage") %>%
  kable_styling(bootstrap_options = "bordered", full_width = FALSE, 
                position = "left", font_size = 12)
```

```{r REF-for-text-below}
# sum of top ten
top_mun_sum <- all_mun %>% 
  slice(1:10) %>% summarise(sum = sum(beneficiarios, na.rm = TRUE))

all_mun_sum <- all_mun %>% 
  summarise(sum = sum(beneficiarios, na.rm = TRUE))
```

> Together, the 10 municipalities with the highest reach (above left) form __`r round(top_mun_sum / all_mun_sum * 100)`%__ of the __`r format(round(sum(u_ben2$beneficiarios)), big.mark = ",")`__ beneficiaries reached. The average coverage of the school-age population in the municipalities where UNICEF is present is __`r round(mean(all_mun$coverage_percent, na.rm = TRUE))`%__. Coverage refers to the percentage of enrolled children (aged 3-17 years) reached by UNICEF. 

&nbsp;

### 4e. Histogram of Coverage

> Below is a histogram of munciipalities where UNICEF is present showing the coverage of enrolled children (aged 3-17). Of note, we have reached 10% or less of the population in __`r sum(all_mun$coverage_percent <= 10, na.rm = TRUE)`__ out of the __`r n_distinct(u_ben2$pcode2)`__ in which we operate. This is in addition to the __`r sum(is.na(all_mun$coverage_percent))`__ where no UNICEF Education activities have occurred. 

```{r PLOT-histogram-of-coverage}
all_mun %>% 
  ggplot(aes(x = coverage_percent)) +
  geom_histogram(binwidth = 10, colour = "black", fill = "white") +
  stat_bin(binwidth = 10, geom = "text", aes(label = ..count..), vjust = -0.5) +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90)) +
  ylab("Number of municipalities") + xlab("Percent of enrolled children (3-17) covered")

```

&nbsp;

## 5. Reports about partners

### 5a. Summary by partner

```{r PLOTLY-stacked-bar-partner, fig.width=9}
# for the totals per partner shows in the geom_text below
stack_text <- act_ben2 %>% 
  group_by(organizacion_implementadora) %>% 
  summarise(total = sum(beneficiarios)) %>% 
  arrange(desc(total))

# vector for ordering of x-axis (partners)
ord_soc <- c("UNICEF", "SOCIO 1", "SOCIO 2", "SOCIO 4", "SOCIO 5", "SOCIO 7", "SOCIO 8",
             "SOCIO 3", "SOCIO 10", "SOCIO 6")

# creating the plot
soc_stack <- act_ben2 %>% 
  group_by(actividad_short, organizacion_implementadora) %>% 
  summarise(beneficiarios = round(sum(beneficiarios))) %>%
  ggplot(aes(x = organizacion_implementadora, 
             y = beneficiarios)) +
  geom_col(aes(fill = actividad_short)) +
  scale_y_continuous(labels = unit_format(unit = "K", scale = 1e-3, sep = ""), 
                     limits = c(0, 400000)) +
  xlab("") + ylab("Beneficiarios") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.key.size = unit(0.1, "cm"),
        legend.position = "bottom",
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8)) +
  scale_fill_discrete(name = "Actividad") +
  geom_text(data = stack_text, aes(y = total + 6000, label = scales::comma(total)), 
            size = 3) +
  scale_x_discrete(limits = ord_soc)

ggplotly(soc_stack, tooltip = c("x", "fill", "y")) %>% 
  layout(legend = list(title = list(text = "<b> organizacion <b>"))) %>% 
  config(displayModeBar = FALSE) %>% 
  layout(showlegend = TRUE, legend = list(font = list(size = 6)))

```

&nbsp;

### 5b. Number of activity types implemented by each partner

```{r TABLE-partner-activity-count}
# this includes CA1.05
act_ben %>% 
  mutate(organizacion_implementadora = 
           str_replace_all(organizacion_implementadora, 
                           "UNICEF - Fondo de las Naciones Unidas para la Infancia",
                           "UNICEF")) %>% 
  rename(partner = organizacion_implementadora) %>% 
  group_by(partner) %>% 
  summarise(act_types = n_distinct(actividad_codigo)) %>% 
  arrange(desc(act_types)) %>% 
  t() %>% 
  pander()

```

&nbsp;

### 5c. Partners' progress over time
*mouse over for details*

```{r line-PLOTLY-partners-progress, fig.width=9}

soc_line <- act_ben2 %>% 
  mutate(mes_solo = as.Date(mes_solo)) %>% 
  group_by(organizacion_implementadora, mes_solo) %>% 
  summarise(cumulative = round(sum(beneficiarios))) %>% 
  mutate(cumulative = cumsum(cumulative)) %>% 
  ggplot(aes(x = mes_solo, y = cumulative, colour = organizacion_implementadora)) +
  geom_line() +
  geom_point(size = 0.7) +
  scale_x_date(date_labels = "%b-%Y", date_breaks = "1 month") +
  scale_y_continuous(labels = unit_format(unit = "K", scale = 1e-3, sep = ""), 
                     limits = c(0, 410000)) +
  xlab("") + ylab("Beneficiarios") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8))

ggplotly(soc_line, tooltip = c("x", "y", "colour")) %>% 
  layout(legend = list(title = list(text = "<b> organizacion <b>"))) %>% 
  config(displayModeBar = FALSE)
  
```

&nbsp;

### 5d. Summary table of partners' achievements

```{r TABLE-partners-summary}
act_ben2 %>% 
  mutate(sex_ben = ifelse(str_detect(desagregacion, "^m"), "male",
                          ifelse(str_detect(desagregacion, "^f"), "female", "no_esp"))) %>% 
  mutate(actividad = actividad_desc) %>% 
  group_by(organizacion_implementadora) %>% 
  summarise(male = sum((beneficiarios[sex_ben == "male"])),
            female = sum((beneficiarios[sex_ben == "female"])),
            sex_ratio = round((male/female), digits = 2),
            beneficiarios = sum(beneficiarios, na.rm = TRUE)) %>% 
  mutate(percent_of_total = round((beneficiarios / sum(beneficiarios) * 100), digits = 2)) %>% 
  left_join(act_ben %>% 
              mutate(organizacion_implementadora = 
              str_replace_all(organizacion_implementadora, 
                           "UNICEF - Fondo de las Naciones Unidas para la Infancia",
                           "UNICEF")) %>% 
              group_by(organizacion_implementadora) %>% 
              summarise(municipalities = n_distinct(pcode2)), 
     by = "organizacion_implementadora") %>% 
  relocate(beneficiarios, .after = organizacion_implementadora) %>% 
  relocate(percent_of_total, .after = beneficiarios) %>% 
  arrange(desc(beneficiarios)) %>% 
  pander(big.mark = ",", style = "rmarkdown")
```

&nbsp;

## 6. Maps and reference table

&nbsp;

### 6a. Maps at municipal level

```{r MAPS-mun-reached-and-percent-reached}

# reading in shapefile
# quiet = TRUE is very important here
pcode2_shape <- st_read("C:/Users/Sean Ng/Documents/R/5W/vnz_adm2_shapefiles/ven_admbnda_adm2_20180502.shp", quiet = TRUE) %>% 
  rename(pcode1 = ADM1_PCODE,
         pcode2 = ADM2_PCODE)

# I think plotly's selector tool is not very good. or maybe the centroids
# of the ven_admbnda dataset is not very good, at least when it comes to centroids. 

# plot for beneficiaries reached by pcode2
mun_ben_plot <- all_mun %>% 
  right_join(pcode2_shape) %>% 
  st_as_sf() %>% 
  ggplot(aes(fill = beneficiarios)) +
  geom_sf(size = 0.01) +
  theme_void() +
  ggtitle("Number of beneficiaries reached by municipality") +
  scale_fill_gradient(trans = trans_reverser("log"), breaks = c(0, 100, 1000, 10000, 80482),
                      labels = comma, na.value = "gray90") +
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = c(0.9, 0.1),
        legend.title = element_blank(),
        plot.title = element_text(size = 12))

# map of percent of enrolled chidren covered by pcode2
mun_percent_plot <- all_mun %>% 
  right_join(pcode2_shape) %>% 
  st_as_sf() %>%
  ggplot(aes(fill = coverage_percent))+
  geom_sf(size = 0.01) +
  theme_void() +
  ggtitle("Percent of enrolled children (3-17) reached by municipality") +
  scale_fill_gradient(labels = comma, trans = "reverse", na.value = "gray90")+
  theme(legend.key.size = unit(0.3, "cm"),
        legend.position = c(0.9, 0.1),
        legend.title = element_blank(), 
        plot.title = element_text(size = 12))

# using patchwork to put them side by side 
mun_ben_plot + mun_percent_plot + plot_layout(ncol = 2)
```

&nbsp;

### 6b. Reference table -- municipal level

__use `UNICEF_present` to filter to municipalities where the Education programme operates__

*CA01.05 Promocion de mensajes claves para la comunidad escolar is not included*

```{r DT-datatable}

# it shows 335, shouldn't there be 336? I bet it's Vargas again
all_mun %>% 
  mutate(UNICEF_present = ifelse(is.na(beneficiarios), FALSE, TRUE)) %>% 
  rename(no_of_activities = act_types) %>% # I feel like you should have done this earlier
  left_join(cen_ref %>% # additional census variables
      group_by(pcode2) %>% 
      summarise(poor_persons = round(sum(ham_2019_xx_poblacion_pobre_por_parroquia)),
                total_pop    = sum(x_2019_poblacion_parroquial_total)) %>% 
      mutate(poverty_incidence = round(poor_persons / total_pop * 100))) %>% 
  select(!pcode2) %>%
  relocate(UNICEF_present) %>% 
  relocate(poverty_incidence, .after = coverage_percent) %>% 
  relocate(no_of_activities, .after = poverty_incidence) %>%
  datatable(filter = "top", options = list(pageLength = 10, scrollX = TRUE)) %>% 
  DT::formatStyle(columns = colnames(.), fontSize = "12pt")
  
```

